---
title: "Association of microbiome vs strange situation in GIMA dataset"
author:
- Kai Xia
date: "`r format(Sys.time(), '%d %B %Y')`"

output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE}
### create data file Microbiome diversity vs Strange situation

#library(xlsx)
library(knitr)
library(tidyr)

#dataset = 'yr1'
if(Sys.info()["sysname"] == 'Windows'){
  dir1 = 'C:/Users/kxia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
} else{
  dir1 = '/Users/kaixia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
}


setwd(dir1)


dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)

dir_beha = sprintf('%s/behaviordata', dir_raw)

dir_ibqr = sprintf('%s/questionnaires', dir_raw)

# if(dataset == 'yr1'){
#   picrustName = '1yr'
# }
# dir_picrust = sprintf('picurst/%spicrust',picrustName)
# L_names = c('L1','L2','L3')
# for (ii in 1:L_names){
#   
# }

datasets = c('neo','yr1')

#########################################################################
### ATTN
### need to use the original version of covariates so that I won't miss the or do we have to?

## neo diversity
for(dataset in datasets){
  
  #dataset = 'neo'
  filein_diversity = sprintf('%s/%s_diversity_covariate.csv',dir_mb,dataset)
  dat1 = read.csv(filein_diversity,header=TRUE,strings=FALSE, check.names = FALSE)
  
  filein1_beha = sprintf('%s/strange_situation_sum.txt',dir_beha)
  data1_beha = read.delim(filein1_beha,header=TRUE,strings=FALSE)
  
  filein2_beha = sprintf('%s/strange_situation_episodes.txt',dir_beha)
  data2_beha = read.delim(filein2_beha,header=TRUE,strings=FALSE)
  
  data_beha = merge(data1_beha,data2_beha,by.x=1,by.y=1)
  
  ## load the IBQR and pycrust 
  filein_ibqr = sprintf('%s/IBQr_17NOV18.csv',dir_ibqr)
  data_ibqr = read.csv(filein_ibqr, header=TRUE,strings=FALSE)[,c(1,3)]
  
  
  ## combine SS and IBQR data
  data_ss_ibqr = merge(data_beha, data_ibqr, by = 1, all = TRUE, sort = FALSE)
  fileout_ss_ibqr = sprintf('%s/data_strange_ibqr_%s.csv', dir_mb, dataset)
  write.csv(data_ss_ibqr, file = fileout_ss_ibqr, row.names = FALSE)
  
  ### create summary results per subject
  mx1 = merge(dat1, data_beha,by.x=1,by.y=1,all.x=TRUE)
  colnames(mx1)[1] = 'SUBID'
  
  ### merge ibqr
  mx1 = merge(mx1, data_ibqr,by.x=1,by.y=1,all.x=TRUE)
  
  
  
  fileout_diversity_beha = sprintf('%s/diversity_strange_%s.csv',dir_mb,dataset)
  write.csv(mx1, file = fileout_diversity_beha, row.names = FALSE)
  
  ### create long format (one record per observation) for linear mixed effect model.
  mx1$SUBID = factor(mx1$SUBID)
  
  
  gathercols <- c("Episode3.1","Episode3.2","Episode3.3")
  
  idx_col = match(gathercols,colnames(mx1))
  mx1_long = gather(mx1, episodes, strange_response, idx_col)
  
  mx1_long$episodes[mx1_long$episodes=="Episode3.1"] <- "1"
  mx1_long$episodes[mx1_long$episodes=="Episode3.2"] <- "2"
  mx1_long$episodes[mx1_long$episodes=="Episode3.3"] <- "3"
  
  
  fileout_diversity_beha = sprintf('%s/diversity_strange_%s_long.csv',dir_mb,dataset)
  
  write.csv(mx1_long, file = fileout_diversity_beha, row.names = FALSE)

}

```






```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE}
### create data file Microbiome vs picrust
### do we run association analysis between diversity to picrust or strange situation to picrust? need Alex input

#library(xlsx)
library(knitr)
library(tidyr)

#dataset = 'yr1'


#dataset = 'yr1'
if(Sys.info()["sysname"] == 'Windows'){
  dir1 = 'C:/Users/kxia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
} else{
  dir1 = '/Users/kaixia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
}


setwd(dir1)


dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)

dir_beha = sprintf('%s/behaviordata', dir_raw)

dir_ibqr = sprintf('%s/questionnaires', dir_raw)



datasets = c('neo','yr1')

## neo diversity
for(dataset in datasets){
  
  # dataset = datasets[1];
  picrustName = dataset
  if(dataset == 'yr1'){
    picrustName = '1yr'
  }
  dir_picrust = sprintf('%s/picrust/%spicrust', dir_raw, picrustName)
  dat1_pi = c()
  
  L_names = c('L1','L2','L3')
  for (ii in 1:length(L_names)){
    file_picrust = sprintf('%s/%s%sKEGG.txt',dir_picrust, picrustName,L_names[ii])
    dat0 = read.delim(file_picrust, header=TRUE, strings=FALSE, skip = 1)
    keggNames = paste(L_names[ii],'.',dat0[,1],sep='')
    dat1 = t(dat0[,-1])
    colnames(dat1) = keggNames
    
    dat1_pi = cbind(dat1_pi, dat1)
  }
  id_names = rownames(dat1_pi)
  ids = gsub('\\.NEO','',id_names)
  ids = gsub('\\.1yr','',ids)
  dat1_pi = data.frame(ID = ids, dat1_pi)
  
  #dataset = 'neo'
  filein_diversity = sprintf('%s/%s_diversity_covariate.csv',dir_mb,dataset)
  dat1 = read.csv(filein_diversity,header=TRUE,strings=FALSE, check.names = FALSE)
  
  mx1 = merge(dat1,dat1_pi,by.x=1, by.y = 1)
  
  fileout_diversity_picrust = sprintf('%s/diversity_picrust_%s.csv',dir_mb,dataset)
  
  write.csv(mx1, file = fileout_diversity_picrust, row.names = FALSE) 

}

```


# Correlation between Strange Situation Outcome, Mask task and IBQ-R

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE}
### create data file Microbiome vs picrust
### do we run association analysis between diversity to picrust or strange situation to picrust? need Alex input

#library(xlsx)
library(knitr)
library(tidyr)

#dataset = 'yr1'


#dataset = 'yr1'
if(Sys.info()["sysname"] == 'Windows'){
  dir1 = 'C:/Users/kxia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
} else{
  dir1 = '/Users/kaixia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
}


setwd(dir1)


dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)

dir_beha = sprintf('%s/behaviordata', dir_raw)

dir_ibqr = sprintf('%s/questionnaires', dir_raw)



dataset = 'neo'
  
  #dataset = 'neo'
  filein_mask = sprintf('%s/data_%s_beha.csv',dir_mb,dataset)
  dat_mask = read.csv(filein_mask,header=TRUE,strings=FALSE, check.names = FALSE)
  
  filein_ss = sprintf('%s/diversity_strange_%s.csv',dir_mb,dataset)
  dat_ss = read.csv(filein_ss,header=TRUE,strings=FALSE, check.names = FALSE)
  
  cvrtNames1 = c('SUBID','MaskAverageScore_FacialFear','MaskAverageScore_VocalDistress'	,'MaskAverageScore_BodilyFear','MaskAverageScore_StartleResponse','MaskAverageScore_EscapeBehavior')
  dat_mask = dat_mask[,cvrtNames1]
  colnames(dat_mask) = gsub('MaskAverageScore_', '', colnames(dat_mask))
  
  #cvrtNames2 = c('SUBID','Summed','Average','Max', 'IBQr_fear')
  cvrtNames2 = c('SUBID','Average', 'IBQr_fear')

  dat_ss = dat_ss[, cvrtNames2]
  
  colnames(dat_ss)[2] = paste('Strange.',colnames(dat_ss)[2],sep='')
  
  mx1 = merge(dat_mask, dat_ss, by.x=1, by.y = 1)

  cor_mat = cor(mx1[,-1], method='spearman', use = "pairwise.complete.obs")
  
  
  print(kable(round(cor_mat,2),caption = sprintf('Correlation matrix of Mask task and strange situation')))

```



# Association analysis for covariates vs strange situation using linear model

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE}
#### run the association analysis for cvrt vs strange situation using linear model
library(knitr)
library(xlsx)

if(Sys.info()["sysname"] == 'Windows'){
  dir1 = 'C:/Users/kxia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
} else{
  dir1 = '/Users/kaixia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
}

setwd(dir1)

source('data_analysis.R')


dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_beha = sprintf('%s/behaviordata', dir_raw)
dir_ibqr = sprintf('%s/questionnaires', dir_raw)
dir_out = '../results'


dataset = 'yr1'
fileout_diversity_beha = sprintf('%s/data_strange_ibqr_%s.csv',dir_mb,dataset)
dat2 = read.csv(fileout_diversity_beha,header=TRUE, check.names = FALSE)
## remove stranger column
dat2 = dat2[,-which(colnames(dat2) == 'Stranger')]


dataName = 'cvrt_vs_strange'

## load cvrt dataset from xlsx file
filein_cvrt_ss = sprintf('%s/GMIAR21dataset_alldemo_17DEC18.xlsx',dir_cvrt)
filein_cvrt_sheetNames = 'GMIAstrange'
data_cvrt_ss = read.xlsx(filein_cvrt_ss, sheetName = filein_cvrt_sheetNames)

## SS
edataNames_ss = c('Summed','Average','Max','Episode3.1','Episode3.2','Episode3.3')

## remove missing:
idx_rm = which(is.na(dat2[,edataNames_ss[1]]))
if(length(idx_rm) > 0){
  mx2 = dat2[-idx_rm,]
} else{
  mx2 = dat2
}

### merge ss data and 
mx2 = merge(mx2, data_cvrt_ss, by = 1, all = TRUE, sort = FALSE)


cdataNames_ss = colnames(data_cvrt_ss)[-c(1:8)]


  dataName = sprintf('cvrt_vs_strange_%s',dataset)

  pairwise_association_test(mx2,cdataNames_ss,edataNames_ss,dataName=dataName,dir_out = dir_out)
  
  edataNames_ibqr = c('IBQr_fear')
  
  ## load cvrt dataset from xlsx file
filein_cvrt_ibqr = sprintf('%s/GMIAR21dataset_alldemo_17DEC18.xlsx',dir_cvrt)
filein_cvrt_sheetNames = 'GMIAIBQr'
data_cvrt_ibqr = read.xlsx(filein_cvrt_ibqr, sheetName = filein_cvrt_sheetNames)

  
 ## remove missing:
idx_rm = which(is.na(dat2[,edataNames_ibqr[1]]))
if(length(idx_rm) > 0){
  mx3 = dat2[-idx_rm,]
} else{
  mx3 = dat2
}
cdataNames_ibqr = colnames(data_cvrt_ibqr)[-c(1:8)]

mx3 = merge(mx3, data_cvrt_ibqr, by = 1 , all = TRUE, sort = FALSE)



  dataName = sprintf('cvrt_vs_IBQr_%s',dataset)

  pairwise_association_test(mx3,cdataNames_ibqr,edataNames_ibqr,dataName=dataName,dir_out = dir_out) 
  



```





# Association analysis for diversity vs strange situation or questionaire using linear model
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE}
#### run the association analysis for diversity vs strange situation or questionaire using linear model


if(Sys.info()["sysname"] == 'Windows'){
  dir1 = 'C:/Users/kxia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
} else{
  dir1 = '/Users/kaixia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
}

setwd(dir1)

source('data_analysis.R')

dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_beha = sprintf('%s/behaviordata', dir_raw)
dir_ibqr = sprintf('%s/questionnaires', dir_raw)
dir_out = '../results'



datasets = c('neo','yr1')
for(dataset in datasets){
  
  fileout_diversity_beha = sprintf('%s/diversity_strange_%s.csv',dir_mb,dataset)
  dat2 = read.csv(fileout_diversity_beha,header=TRUE, check.names = FALSE)
  dataName = 'diversity_vs_strange'
  #dir_out = 'output'
  
  
  ## save response data
  edataNames = c('Summed','Average','Max','Episode3.1','Episode3.2','Episode3.3','IBQr_fear')
  
  # ## remove missing:
  # idx_rm = which(is.na(dat2[,edataNames[1]]))
  # if(length(idx_rm) > 0){
  #   mx2 = dat2[-idx_rm,]
  # }
  mx2 = dat2
  
  
  cdataNames = c('chao1','observed_otus','PD_whole_tree','shannon'
                 ,'wunifrac.PC.1','wunifrac.PC.2','wunifrac.PC.3','wunifrac.PC.4')
  
  
  dataName = sprintf('strange_vs_diversity_%s',dataset)

  pairwise_association_test(mx2,cdataNames,edataNames,dataName=dataName,dir_out = dir_out)



}

```


# Association analysis for diversity vs strange situation or questionaire using linear mixed effect model
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE}
#### run the association analysis for diversity vs strange situation or questionaire using linear model
library(lme4)
library(knitr)

if(Sys.info()["sysname"] == 'Windows'){
  dir1 = 'C:/Users/kxia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
} else{
  dir1 = '/Users/kaixia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
}

setwd(dir1)

source('data_analysis.R')

dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_beha = sprintf('%s/behaviordata', dir_raw)
dir_ibqr = sprintf('%s/questionnaires', dir_raw)
dir_out = '../results'

randomName = 'SUBID'
cvrtNames = 'episodes'


datasets = c('neo','yr1')

for(dataset in datasets){
  # dataset = datasets[1];
  fileout_diversity_beha = sprintf('%s/diversity_strange_%s_long.csv',dir_mb,dataset)
  dat2 = read.csv(fileout_diversity_beha,header=TRUE, check.names = FALSE)
  dataName = sprintf('diversity_vs_strange_combine_%s',dataset)

  ## save response data
  edataNames = c('strange_response')
  
  mx2 = dat2
  
  mx2$episodes = as.factor(mx2$episodes)
  
  cdataNames = c('chao1','observed_otus','PD_whole_tree','shannon'
                 ,'wunifrac.PC.1','wunifrac.PC.2','wunifrac.PC.3','wunifrac.PC.4')
  
  
  pairwise_association_test_lmm (mx2,cdataNames,edataNames, randomName = randomName, cvrtNames=cvrtNames
								, dataName=dataName,dir_out = dir_out)


}

```




# Association analysis for diversity vs picrust using linear model
```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE}
#### run the association analysis for diversity vs strange situation or questionaire using linear model
library(knitr)

if(Sys.info()["sysname"] == 'Windows'){
  dir1 = 'C:/Users/kxia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
} else{
  dir1 = '/Users/kaixia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
}

setwd(dir1)

source('data_analysis.R')

dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt',dir_raw)
dir_beha = sprintf('%s/behaviordata', dir_raw)
dir_ibqr = sprintf('%s/questionnaires', dir_raw)
dir_out = '../results'



datasets = c('neo','yr1')
for(dataset in datasets){
  # dataset = datasets[1];
  fileout_diversity_beha = sprintf('%s/diversity_picrust_%s.csv',dir_mb,dataset)
  dat2 = read.csv(fileout_diversity_beha,header=TRUE, check.names = FALSE)
  #dir_out = 'output'
  
  
  ## save response data
  keggNames = c('L1','L2','L3')
  
  for(keggName in keggNames){
    idx = grep(keggName,colnames(dat2))
    # for(keggName in keggNames){
    #   idx = c(idx,grep(keggName,colnames(dat2)))
    # }
    cdataNames = colnames(dat2)[idx]
    
    # ## remove missing:
    # idx_rm = which(is.na(dat2[,edataNames[1]]))
    # if(length(idx_rm) > 0){
    #   mx2 = dat2[-idx_rm,]
    # }
    mx2 = dat2
    
    dataName = sprintf('diversity_vs_picrust_%s_%s',keggName,dataset)
  
    edataNames = c('wunifrac.PC.1','wunifrac.PC.2','wunifrac.PC.3','wunifrac.PC.4')
    
    
    pairwise_association_test (mx2,cdataNames,edataNames
  								, dataName=dataName,dir_out = dir_out)
  
  }
}

```





## Microbiome alpha diversity difference (yr1 vs neo) vs SS, IBQ-R

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE}
### neo vs brain volume data

library(knitr)
dataset0 = 'neo'
dataset1 = 'yr1'
dataset = 'diff'

if(Sys.info()["sysname"] == 'Windows'){
  dir1 = 'C:/Users/kxia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
} else{
  dir1 = '/Users/kaixia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
}

setwd(dir1)

source('data_analysis.R')

dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt', dir_raw)
#dir_brain = sprintf('%s/braindata', dir_raw)
dir_out = '../results'


div_Names = c("chao1", "observed_otus", "PD_whole_tree", "shannon")

##load the SS IBQ-R data
fileout_diversity_beha = sprintf('%s/diversity_strange_%s.csv',dir_mb,dataset1)
dat_ss = read.csv(fileout_diversity_beha,header=TRUE, check.names = FALSE)

## remove the div in dat_ss
dat_ss = dat_ss[, -which(colnames(dat_ss) %in% div_Names)]


#dataName = 'div_diff_vs_strange'

edataNames = c('Summed','Average','Max','Episode3.1','Episode3.2','Episode3.3','IBQr_fear')




filein0_diversity = sprintf('%s/%s_diversity_covariate.csv',dir_mb,dataset0)
dat0 = read.csv(filein0_diversity,header=TRUE,strings=FALSE,check.names=FALSE)

filein1_diversity = sprintf('%s/%s_diversity_covariate.csv',dir_mb,dataset1)
dat1 = read.csv(filein1_diversity,header=TRUE,strings=FALSE,check.names=FALSE)





colNames = c("SUBID", div_Names)

dat0 = dat0[,colNames]
dat1 = dat1[,colNames]
mx1 = merge(dat0,dat1,by.x=1,by.y=1)
#div_Names = c('chao1','observed.species','PD.whole','shannon')
for( div_Name in div_Names){
  name_x = paste(div_Name,'.x',sep='')
  name_y = paste(div_Name,'.y',sep='')
  mx1[,div_Name] = mx1[,name_y] - mx1[,name_x]
}



mch1 = match(dat_ss[,1], mx1[,1])

mx2 = merge(dat_ss, mx1,by.x=1,by.y=1)



### running regression between diversity and 
cdataNames = colNames[-1]

  dataName = sprintf('div_diff_vs_strange_%s',dataset1)

  pairwise_association_test(mx2,cdataNames,edataNames,dataName=dataName,dir_out = dir_out)
  
```





## Microbiome alpha diversity difference (yr1 vs neo) vs SS, IBQ-R using linear mixed effect model

```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE}
### neo vs brain volume data

library(knitr)
library(lme4)

dataset0 = 'neo'
dataset1 = 'yr1'
dataset = 'diff'

if(Sys.info()["sysname"] == 'Windows'){
  dir1 = 'C:/Users/kxia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
} else{
  dir1 = '/Users/kaixia/OneDrive - University of North Carolina at Chapel Hill/github/gmia/scripts'
}

setwd(dir1)

source('data_analysis.R')

dir_mb = '../processed_data'
dir_raw = '../raw_data'
dir_cvrt = sprintf('%s/cvrt', dir_raw)
#dir_brain = sprintf('%s/braindata', dir_raw)
dir_out = '../results'


div_Names = c("chao1", "observed_otus", "PD_whole_tree", "shannon")

##load the SS IBQ-R data
fileout_diversity_beha = sprintf('%s/diversity_strange_%s_long.csv',dir_mb,dataset1)
dat_ss = read.csv(fileout_diversity_beha,header=TRUE, check.names = FALSE)

## remove the div in dat_ss
dat_ss = dat_ss[, -which(colnames(dat_ss) %in% div_Names)]


#dataName = 'div_diff_vs_strange'

edataNames = c('strange_response')
randomName = 'SUBID'
cvrtNames = 'episodes'



filein0_diversity = sprintf('%s/%s_diversity_covariate.csv',dir_mb,dataset0)
dat0 = read.csv(filein0_diversity,header=TRUE,strings=FALSE,check.names=FALSE)

filein1_diversity = sprintf('%s/%s_diversity_covariate.csv',dir_mb,dataset1)
dat1 = read.csv(filein1_diversity,header=TRUE,strings=FALSE,check.names=FALSE)


colNames = c("SUBID", div_Names)

dat0 = dat0[,colNames]
dat1 = dat1[,colNames]
mx1 = merge(dat0,dat1,by.x=1,by.y=1)
#div_Names = c('chao1','observed.species','PD.whole','shannon')
for( div_Name in div_Names){
  name_x = paste(div_Name,'.x',sep='')
  name_y = paste(div_Name,'.y',sep='')
  mx1[,div_Name] = mx1[,name_y] - mx1[,name_x]
}



mch1 = match(dat_ss[,1], mx1[,1])

mx2 = merge(dat_ss, mx1,by.x=1,by.y=1)
mx2$episodes = as.factor(mx2$episodes)


### running regression between diversity and 
cdataNames = colNames[-1]

  dataName = sprintf('div_diff_vs_strange_combine_%s',dataset1)

  pairwise_association_test_lmm (mx2,cdataNames,edataNames, randomName = randomName, cvrtNames=cvrtNames
								, dataName=dataName,dir_out = dir_out)
  
```




  
  
  